---
title: "lab_02"
author: "Derek Willis"
date: "2026-02-18"
output: html_document
name: "Riley Bomgardner"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1. 32 departments have offered "Climate" courses, with the Atmospheric and Oceanic Science department offering the most, at 92 courses.**

```{r}
climate_courses <- umd_courses |>
  filter(str_detect(title, "Climate"), seats >= 1) |>
  group_by(department) |>
  summarize(num_courses = n()) |>
  arrange(desc(num_courses))
```

**Q2.** Of the departments listed in A1, which one is the most interesting to you, and why? **A2. The journalism department offers two courses, which is interesting. To me, climate can mean different things, so I wonder if this is more about political climate or how to handle that structure as you get into your career.**

**Q3** Which departments offered the most courses in the Fall 2024 term? And which departments in that term have at least one syllabus posted? Answer both of those questions and then compare the top five departments - how are they different?

**A3. The departments that offered the most courses in Fall 2024 were Business and Management, School of Music, Information Studies, English, and Teaching and Learning, Policy and Leadership. Only 134 of the 196 departments for the 2024 Fall semester have at least one syllabus for its classes. Economics gets pushed into second for most syllabi, but that's the only big difference among the top five.**

```{r}
# Top departments by total courses in Fall 2024
fall_2024_courses <- umd_courses |>
  filter(term == 202408) |>
  group_by(department) |>
  summarize(
    total_courses = n(),
    courses_with_syllabus = sum(syllabus_count >= 1)
  ) |>
  arrange(desc(total_courses))

# Departments with at least one syllabus posted
fall_2024_syllabus <- fall_2024_courses |>
  filter(courses_with_syllabus >= 1) |>
  arrange(desc(courses_with_syllabus))
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as a new dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name, and we're not looking in payee_name). Write a sentence describing the pattern of spending over time. **A4. Spending by Wes Moore's committee seems to increase every month, with some outliers, but as the election nears, the last couple months decrease very drastically with their spending.**

```{r}
maryland_expenses |>
  mutate(
    expenditure_date = mdy(expenditure_date),
    month = floor_date(expenditure_date, "month")
  ) |>
  filter(committee_name == "Moore  Wes For Maryland") |>
  group_by(month) |>
  summarize(total_amount = sum(amount, na.rm = TRUE)) |>
  arrange(month)
```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for all others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5. Around 78 million dollars came from in-state accounts, while about 52 million accounts were out of state.**

```{r}
maryland_expenses |>
  mutate(
    expenditure_date = mdy(expenditure_date),
    spending_location = case_when(
      str_detect(address, " Maryland ") ~ "In-state",
      .default = "Out of state"
    )
  ) |>
  group_by(spending_location) |>
  summarize(total_amount = sum(amount, na.rm = TRUE)) |>
  arrange(desc(total_amount))
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6.Which payees sent the most total money to Moore's committee after standardizing names to uppercase? MOORE MILLER FOR MARYLAND SLATE sent the most money to support Moore's campaign, with 4,867,000 dollars.**

```{r}
maryland_expenses |>
  mutate(
    expenditure_date = mdy(expenditure_date),
    payee_upper = str_to_upper(payee_name)
  ) |>
  filter(committee_name == "Moore  Wes For Maryland") |>
  group_by(payee_upper) |>
  summarize(total_amount = sum(amount)) |>
  arrange(desc(total_amount))
```
