---
title: "week16_recap"
author: "Daniel Trielli"
date: "2026-05-07"
output: html_document
name: "Riley Bomgardner"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(tigris)
library(tidycensus)
library(corrr)
```

## Load Data

[USAFacts compiled data on inspections on child care facilities throughout the country](https://usafacts.org/data-projects/childcare-inspections) - specifically, how many of those facilities are overdue for inspection. Let's take a look on the data they have on Maryland.

First, here the data we need to load:

```{r}
childcare_inspections <- read_csv('data/childcare_inspections.csv')
childcare_inspections$fips <- as.character(childcare_inspections$fips)
```

#### **Q1** The childcare data does not have the names of the counties. Join them with the md_counties data, that can be loaded in the block below. Then inspect the dataframe you created with this join and answer this: there are two columns with county names in the new dataframe. Which one is the best to be used? (hint: it has to do with homonyms)

```{r}
# This loads the county data
md_counties <- counties() |>  filter(STATEFP == "24")
```

```{r}
# Join the childcare data with the county data 
md_childcare <- childcare_inspections |>
  left_join(md_counties, join_by(fips == GEOID))
```

##### A1: The NAMELSAD column should be used because it lists Baltimore city and Baltimore County. The NAME column just has two entries for Baltimore.

------------------------------------------------------------------------

#### **Q2** Create bar chart with the percentages of overdue childcare facility inspections per county to answer (You'll need to create a column for that calculation first). What is the county with the largest percentage of facilities with overdue inspections?

```{r}
# Calculate the percentage of overdue inspections per county

md_childcare <- md_childcare |>
  mutate(percent_overdue = overdue/total * 100)

# Create the chart with ggplot
md_childcare |>
  ggplot() +
  geom_bar(aes(x = NAMELSAD, weight = percent_overdue)) +
  labs(
    title = "Montgomery, Harford, and PG Counties top highest percent of overdue childcare facility inspections",
    x = "County",
    y = "Rate of Overdue Inspections"
  )

```

##### **A2: Montgomery County has the highest rate of overdue childcare facility inspections.**

------------------------------------------------------------------------

#### **Q3** Next, we want to see if the number of child care facilities - with overdue inspections or otherwise - is appropriate for each county. So we're going to calculate the rate of children per facility per county. To do that, you'll first need Census data to get the population of children under 5 for each county. Which county has the biggest rate of children per facility (i.e. they could use more facilities)?

```{r}
# First, grab data from the ACS. The ACS variable for children age under 5 is B06001_002 We want 2022 data.

under_5 <- get_acs(
  geography = "county",
  variables = c(medincome = "B06001_002"),
  year = 2022) |>
  mutate(GEOID = as.character(GEOID))

```

```{r}
# Rename the column name 'estimate', which has our population data, with something more useful, like 'children_pop'

under_5 <- under_5 |>
  rename(children_pop = estimate)

```

```{r}
# Join the dataframes

under_5_at_childcare <- md_childcare |>
  left_join(under_5, join_by(fips == GEOID))

# Calculate the rates for children per facility

under_5_at_childcare <- under_5_at_childcare |>
  mutate(child_per_fac = children_pop/total)

```

##### **A3: Garrett County has the highest rate at 87.5, which means they could definitely use more facilities, but most likely lack the resources and funding in the area.**

------------------------------------------------------------------------

#### **Q4** Make a map of the results for question 3. Are there any areas of the state that are more lacking in childcare facilities?

```{r}
library(sf)

under_5_at_childcare_sf <- st_as_sf(under_5_at_childcare)

under_5_at_childcare_sf |>
  ggplot(aes(fill = child_per_fac)) +
  geom_sf() +
  scale_fill_viridis_c() +
  labs(
    title = "Children per childcare facility in Maryland counties",
    fill = "Children per facility"
  )
```

##### **A4: The pan handle, or the more northwest, mountainous counties have less childcare facilities and more children at each facility. The Eastern Shore is somewhat similar, but not as bad. Is this because they are more rural areas? **

------------------------------------------------------------------------

#### **Q5** Now let's explore if we can find patterns in under-served areas, particularly when it comes to median income of the population. First, join the childcare data median income data from the Census. Then, create a scatterplot and perform a cor.test(), and answer: is there is any correlation between income and rate of children per facilities? What makes you conclude that?

```{r}
# Again, grab data from the ACS. The ACS variable for median income is B19013_001. We want 2022 data again.

median_income <- get_acs(
  geography = "county",
  variables = c(median_income = "B19013_001"),
  year = 2022) |>
  mutate(GEOID = as.character(GEOID))
```

```{r}
# Rename the column name 'estimate', which has our income data, with something more useful, like 'median_income'

median_income <- median_income |>
  rename(median_income = estimate)

```

```{r}
# Join the dataframes

median_income_join <- under_5_at_childcare |>
  left_join(median_income, join_by(fips == GEOID))

```

```{r}
# Create a scatterplot with median income and rate of children per facility

median_income_join |>
  ggplot(aes(x = median_income, y = child_per_fac)) +
  geom_point() +
  labs(
    title = "Not a great correlation between median income and children per facility",
    x = "Median Income",
    y = "Children per facility"
  )

```

```{r}
# Create a correlation test.

cor.test(
  median_income_join$median_income,
  median_income_join$child_per_fac
)
```

##### A5: There is a very, very weak negative correlation between the two variables. For one, the cor test shows that the p-value is over 0.05, which means this correlation is not statistically significant and therefore you cannot say that median income influences the amount of childcare facilities in a county or the number of children that attend each one.

------------------------------------------------------------------------

#### **Q6** Finally, let's do some text analysis. We another dataset, with inspection reports from informal child care providers in the state. This dataset has an informative column of initial findings: a column that describes issues found in these providers, if any. Your goal is to create bigrams for this column and find common issues that are mentioned in these findings. And then answer: what kind of issues listed there you think are interesting? They don't have the most frequent.

```{r}
# Read the reports data

library(rvest)
library(tidytext)

reports <- read_csv('data/childcare-informal-provider-inspections.csv') |> clean_names()
```

```{r}
# Create a stopword dictionary (feel free to include more words)

data("stop_words")

stop_words <- stop_words |>
  add_row(word = "childcare") |>
  add_row(word = "child") |>
  add_row(word = "care")

```

```{r}
# Unnest tokens into bigrams, removing stop words and repasting bigrams, list bigrams

reports |>
  unnest_tokens(bigram, initial_findings, token = "ngrams", n = 2) |>
  separate(bigram, c("word1", "word2"), sep = " ") |>
  filter(!word1 %in% stop_words$word,
         !word2 %in% stop_words$word) |>
  unite(bigram, word1, word2, sep = " ") |>
  count(bigram, sort = TRUE)

```

##### A6: The most common bigrams are "emergency bag," "ertg missing," and "cleaning products." One that stuck out to me was "knife draw." I know what its referring to, which is easily accessible knives or other sharp kitchen items that could potentially be accessible to children at a care facility. But the phrase by itself is kind of silly. It reminds me of, "Who brings a knife to a gun fight?"
