---
title: "Group 1 - CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban"
author: "Riley Bomgardner, Matt Germack"
date: "March 11, 2026"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban](https://www.wral.com/in-session-campaign-donations/17716104/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
library(stringr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

nc_donor_data <- read.csv("data/nc_donors_2018.csv", stringsAsFactors = FALSE)
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [A WRAL News analysis of campaign finance filings also turned up $16,000 in donations to legislative campaigns during session from 15 PACs with a lobbyist. Each would appear to be a violation of state fundraising rules, but spokespeople for most of the PACs, or the companies they represent, said there was a simple explanation: The checks were written before session.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Beer distributors also gave during session. Mark Craig, the president of distributor R.H. Barringer, put more than $30,000 into state legislative campaigns during the six-week period, including the maximum $5,200 to Berger. He was the top individual donor to legislative campaigns during session.]
* **Analysis summary**: [We were able to confirm our finding. Using a string detect, we were able to take out all of Mark Craig's donations to NC legilators, along with the maximum donation of 5200. It's pretty straightforward, but it's still able to be confirmed.]

```{r}
# Put code to reverse engineer sentence here
nc_donor_data |>
  filter(str_detect(name_clean, "MARK R CRAIG")) |>
  summarise(total = sum(amount), max = max(amount))
# Display results of code below this codeblock
#total = 33,200.  max = 5200
```

### Sentence 3

* **Sentence text**: [Add in donations to the two main political parties and to a handful of other large fundraising vehicles for state legislative races, and the in-session giving topped $3.4 million. Gov. Roy Cooper's main campaign account brought in another $407,000 during session.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [This year's legislative session also saw significant donations from hog farmers and beer distributors, both industries with important legislation before the General Assembly. Senate President Pro Tem Phil Berger's campaign alone raised nearly $222,000 during session. Twenty-three people gave him the maximum allowed: $5,200.]
* **Analysis summary**: [Our analysis is correct. In order to do this question, constraints on date needed to align with the time that the six week session started and ended. The total amount donated to Berger during that time was $221925, which is roughly $220,000, which is what the article said. This code also outlines the 23 donors who gave $5200 as their donation. I also added in some code about the hog farmers and beer industry, but to me their donations don't seem that significant.]

```{r}
# Put code to reverse engineer sentence here

full_session <- subset(nc_donor_data, date_clean >= "2018-05-16" & date_clean <= "2018-06-29")

#Berger
berger <- subset(full_session, committee_name == "PHILIP E BERGER COMM")

#Gives full amount of campaign donations during session
sum(berger$amount, na.rm = TRUE)

#gives number of donors to Berger
nrow(subset(berger, amount == 5200))

#hog farmers donations
hog <- subset(full_session, grepl("pork|smithfield|farm bureau", name, ignore.case = TRUE))

sum(hog$amount, na.rm = TRUE)

#beer distributors
beer <- subset(full_session, grepl("beer & wine wholesale", name, ignore.case = TRUE))

sum(beer$amount, na.rm = TRUE)
  
# Display results of code below this codeblock

#Total amount = 221,925
# of donors = 23
#money donated to berger from hog farmers = 16,000
#money donated to berger from beer = $2500

```

### Sentence 5

* **Sentence text**: [There was a major bump during the first week of session this year for the state parties, but that's because three large donors – LinkedIn co-founder Reid Hoffman, Durham investor Greg Lindberg and Lindberg's Eli Research company – donated $1.3 million that week to the state Republican and Democrat parties.]
* **Analysis summary**: [This one was a little more challenging. I had to change the date format so that the computer could figure out when the first week of the session was. Using the donors names was easy, along with what columns I wanted shown. The number came out of 1.25 million, which was most likely rounded up in the article. This code also breaks down how much each of the three donors gave, with Reid Hoffman and Greg Lindberg donated 500,000 dollars in the first week.]

```{r}
# Put code to reverse engineer sentence here

nc_donor_data$amount <- as.numeric(gsub("[$,]", "", trimws(nc_donor_data$amount)))

nc_donor_data$date_clean <- as.Date(nc_donor_data[["date"]], "%m/%d/%y")
 
big_donors <- c("Reid Garrett Hoffman", "Greg Lindberg", "ELI Research LLC")
 
first_week <- subset(nc_donor_data,
                     date_clean >= "2018-05-16" &
                     date_clean <= "2018-05-22" &
                     name %in% big_donors)
 
first_week[, c("date_clean", "name", "committee_name", "amount")]
 
sum(first_week$amount, na.rm = TRUE)

# Display results of code below this codeblock

#total amount for first week = 1,250,000

```

-30-