---
title: "Reverse Engineering Project"
author: "Riley Bomgardner, Adam Lipman, Logan Beebe"
date: "2026-04-27"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Data Link: <https://www.kaggle.com/datasets/andrewmvd/board-games?resource=download>

## Load libraries

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
```

## Loading Data

```{r}
board_games_data <- read.csv2("data/bgg_dataset.csv") |> clean_names()
```

```{r}
board_games_data
```

## Basic Analysis

There are 20,343 rows and 14 columns in the dataset.

We do not have any price data included in this dataset so we will need joins to other datasets to address that. We also need information about attention deficits and ages that play the most. The column for game mechanics in our data, called mechanics, has inconsistent formatting for separate features, using slashes, commas or spaces to divide them all. They are in the same row which makes it more challenging.

Our dataset does not have any timeframe limitations but some older games like marbles have rough approximates that use negative values. Those values are not specified to be time since the present or if it is in a BCE format. The last year with data is 2022, removing any conclusions from the previous three years. The play_time is not the actual length needed to play the game but instead it is biased data from the game's creators. The reviews for the games do not have sources, making it unclear if they are from one source or combined across multiple, which could have different rating criteria.

We plan to research the needed data categories and try to add that for our future deliverables.

```{r}
nrow(board_games_data)
```

Does board game popularity correlate with the number of exclusive game mechanics within it?

Findings: There is a slight correlation, but it's negative (-0.37). This means that a board game with a top bgg_rank (1) will most likely have more mechanics listed than a board game listed in the ranking list at 20,000, and the decrease in mechanic count will decrease as the rank decreases. So, yes, board game ranking, or popularity, does correlate with game mechanics listed.

The correlation test also confirms that this information is statistically significant due to a high t-value and a p-value that's nearly equivalent to 0.

Does a higher amount of mechanics also mean that the complexity rate will be higher? Are the more popular board games ones that are more complex or harder to complete?

```{r}

board_games_data_2 <- board_games_data |>
  mutate(mechanic_count = ifelse(
    is.na(mechanics) | mechanics == "",
    0,
    str_count(mechanics, ",") + 1
  ))

board_games_data_2 |>
  select(name, mechanics, mechanic_count, bgg_rank)

cor.test(board_games_data_2$bgg_rank, board_games_data_2$mechanic_count)
```



What board games are the most optimized for kids with attention disorders? In other words, what ADHD-friendly board games, or ones with a shorter playing time, take into account shorter attention spans of kids with ADHD? The max amount of playing time would be around 30 minutes. This would also be filtered with a minimum age to play. Strategy game mechanics are also considered.

Findings: The top game that would be considered AHDH-friendly under these guidelines is TerroriXico. It's a Mexican board game where players work together to capture monsters by using different cards throughout the game. The player with the most victory points wins. This seems like a game that will definitely keep a person's attention on what's happening in front of them, and it doesn't sound boring at all. Other top board games include Flip and Fish, Dodos and Dinos, and Animator vs. Animation. 

```{r}
library(dplyr)
library(stringr)
library(readr)

# 1. Convert only what we need
board_games_data_2 <- board_games_data_2 |>
  mutate(
    play_time = as.numeric(`play_time`),
    min_age = as.numeric(`min_age`),
    rating_average = as.numeric(str_replace_all(`rating_average`, ",", "."))
  )

# 3. Filter: play time and min age only
results <- board_games_data_2 |>
  filter(
    play_time > 0,
    play_time <= 30,
    min_age   > 0,
    min_age   <= 10
  ) |>
  arrange(desc(rating_average)) |>
  select(name, min_age, play_time, rating_average, domains, mechanics)

print(results)
```