---
title: "week7_recap"
author: "Daniel Trielli"
date: "2026-03-11"
output: html_document
name: "Riley Bomgardner"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load Data

Run the codeblock below to load the data.

```{r}
earthquakes <- read_csv('https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv')

#Setting time column as datetime
earthquakes <- earthquakes |> mutate(time = as_datetime(time))
```

#### Answer the questions below

Most questions have a code block and a space for an answer below. Write the code you think is necessary and, in the answer space, write out what you did and what was the result.

------------------------------------------------------------------------

#### **Q1** Do you notice anything different with this read_csv()? Particularly in the file name?

**A1: Yes, instead of an actual file being used for our data, a URL is used. We're getting data from the internet, not in a zip file that was given to us.**

------------------------------------------------------------------------

#### **Q2** How many records there are there in this dataset? What do they mean and what useful information we can gather from it, looking at the columns?

**A2: There are 10,309 records. Each row represents a new earthquake somewhere around the world, with other measures like depth, magnitude and location. This is a large number of earthquakes in a span of 30 days.**

------------------------------------------------------------------------

#### **Q3** How do I reorganize this list to see the ones that are the deepest first? What is the depth that shows up for the deepest one, and its magnitude?

```{r}
earthquakes |>
  arrange(desc(depth))
```

**A3: The deepest earthquake is in Fiji on February 14, with a depth of 667.197. Its magnitude was a 4.30.**

------------------------------------------------------------------------

#### **Q4** I just want to see the earthquakes with a magnitude larger than 6. How do I do that? And how many are there that fit this criteria?

```{r}
earthquakes |>
  filter(mag > 6)

```

**A4: 9 rows with magnitude larger than 6.**

------------------------------------------------------------------------

#### **Q5** What about if I want to see earthquakes that have both a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [both]{.underline} these criteria?

```{r}
earthquakes |>
  filter(mag > 6 & depth < 20)

```

**A5: 4 fit both**

------------------------------------------------------------------------

#### **Q6** What about if I want to see earthquakes that either have a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [either]{.underline} these criteria?

```{r}
earthquakes |>
  filter(mag > 6 | depth < 20)

```

**A6: 8,250 rows fit either.**

------------------------------------------------------------------------

#### **Q7** I'm interested in finding earthquakes that took place in Alaska. Which column should I look at? How do I use it to find all the earthquakes in Alaska? How many earthquakes took place there?

```{r}
earthquakes |>
  filter(str_detect(place, ", Alaska"))

```

**A7: 3326 earthquakes have taken place in Alaska over the last 30 days.**

------------------------------------------------------------------------

#### **Q8** I notice that there is a column called 'type', that seems to have different kinds of tremors. What is the best way to find what are all the possible types of tremors, and counting how many of each there are in this data set? What are the first two most common types of tremors in this data set?

```{r}
earthquakes |>
  group_by(type) |>
  summarize(count = n()) |>
  arrange(desc(count))
```

**A8: 10124 earthquakes, 69 explosions, 15 ice quakes, 1 mining explosion, 98 quarry blasts. The two most common tremors are earthquakes and quarry blasts.**

------------------------------------------------------------------------

#### **Q9** How do I calculate the average depth of each type of tremor? What is the average depth of the earthquake type in this data set? Is there anything that seems unusual or surprising in this finding?

```{r}
earthquakes |>
  group_by(type) |>
  summarize(average_depth = mean(depth)) |>
  arrange(desc(average_depth))
```

**A9: Average depth of an earthquake is 21.867. Quarry blasts and explosions are above ground. Why did only one mining explosion set off an earthquake?**

------------------------------------------------------------------------

#### **Q10** I'm interested, in the future, to see the hours in which earthquakes happen. How can I extract the hour from the time column?

```{r}
earthquakes_with_hour <- earthquakes |>
  mutate(hour = hour(time))

```

**A10: By using an hour function.**

------------------------------------------------------------------------

#### **Q11** I want to make a note of all the records in this data set that I consider serious. For my purposes, I'm thinking that all tremors that have a magnitude that is larger than 3 are serious. How do I automatically create a new column making that distinction?

```{r}
earthquakes |>
  mutate
case_when
```

**A11:**

------------------------------------------------------------------------

#### **Q12** I have no idea how earthquakes work and I'm interested in seeing if there is a particular time of day in which serious earthquakes happen. How can I see that condensed in a table with all the hours in a day and all the serious earthquakes in each hour? What is the hour with fewer serious earthquakes and the one with the most serious earthquakes?

```{r}
filter
group_by()
```

**A12**:

------------------------------------------------------------------------

#### EXTRA Q What other questions are you interested in and how would you ask them in R?

```{r}

```

```{r}

```

```{r}

```
